Formalizing Stack Safety as a Security Property

### 2023-01-20 submission5/core.tex
Stray S restored
Editing through key ideas

Stack safety is challenging to specify in full generality, because the stack is fundamentally
an architecture-specific construct. (Contrast Acevedo de Amorim et al.'s \cite{} work on heap safety,
which is broadly consistent across architectures and therefore amenable to a very general approach.)
The desired behavior of the stack varies based on ISA, calling convention,
and the features of whichever programming language originated the code.
For instance, the current state-of-the-art stack-safety specification from Georges et al.
\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
a callee should always return to its caller. But, in the presence of tail-call elimination, a
return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
must be accompanied by tweaks to the model.

@@

Security properties tend to be conservative, identifying violations that would be
harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
and of compatibility with existing codebases. In the motivating case for this work,
Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
A set of software-defined rules, termed a micro-policy, determines the tags for operation
results based on the tags of their inputs---or, if the input tags are in a dangerous
combination, it may halt the machine (a ``failstop.'')

Roessler and DeHon find that the obvious way to enforce stack safety in this setting
is inefficient when allocating and clearing stack frames, because each word must be initialized
with a tag corresponding to the (depth of the) current function activation. They propose
{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
when a function writes to memory. This means that such writes cannot be protected. Instead
the micro-policy enforces that memory be read by the same fuction that wrote it. This
lazy enforcement is incompatible with existing models of stack safety, particularly the
integrity side, because an attacker can modify another function's data. But the owner of
the data will never see the changes, so it may still be secure in some sense. We aim to elaborate
that sense, and to test whether the micro-policy fulfills it.
This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
machine states that make claims about the future observable behavior of execution from those states.

@@

The paper is structured as follows. In \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. Then in \cref{sec:props} we will introduce the machine model that we test, and give
its {\em security semantics}, described in brief below.
%In \cref{sec:props} we give abbreviated
%formal statements of the properties, while
The full formalism is found in \cref{app:props}.
\Cref{sec:testing} describes the testing framework used to test it.
\Cref{sec:relwork,sec:future} discuss related and future work.
The remainder of this section introduces ideas that will be important throughout.

A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
\emph{values}. Each value \(v\) contains a \emph{payload} word, written \(|v|\).
The details of value structure depend on the specific machine being modelled;
intuitively, the payload represents the part of the value that is relevant to
the behavior of the basic machine, while the rest of the value may contain
information relevant to a hardware enforcement mechanism (such as a tag).
We write \(\mach[\component]\) to denote the value of \(\mach\) at
\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).

\paragraph*{Security Semantics}

@@

The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. Those that we explore in detail in \cref{sec:examples}
are shown in white: calls, returns, and the allocation and deallocation of local memory.
In the full formalism (\cref{app:full}) and the machine that we test,
we cover additional operations add parameters to our existing ones, to support
various forms of memory sharing and tail-call elimination. These are shown in light gray.
We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
based on pointer provenance in \cref{app:ptr}, represented by the dark grey
operations, but we do not test this.

@@

\begin{table}
  \begin{tabular}{| l | l |}
    \hline
    \(\psi \in \Psi\) & Parameters \\
    \hline
    \example \(\mathbf{call}\) & Target address \\
    \example & Argument registers \\
    \testing & Stack argument offset \& size \\
    \example \(\mathbf{return}\) & \\
    \example \(\mathbf{alloc}\) & Offset \& size \\
    \testing & Public flag \\
    \example \(\mathbf{dealloc}\) & Offset \& size \\
    \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
    \hline
    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
    \hline
    \theory \(\mathbf{promote}\) & Register, offset \& size \\
    \theory \(\mathbf{propagate}\) & Source register/address \\
    \theory & Destination register/address \\
    \theory \(\mathbf{clear}\) & Target register/address \\
    \hline
  \end{tabular}
  \label{tab:psi}
\end{table}

\vspace{\belowdisplayskip}

@@

\paragraph*{Security Classes}

The security principals in our account of stack safety are function
activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
it may be private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)). The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.

\paragraph*{Events, Steps, and Traces}

@@

Events are ranged over by \(\obs\), and include the special silent event
\(\tau\). Traces are non-empty, potentially infinite (coinductive) series of
events, and traces are {\em similar} \(\obsT_1 \simeq \obsT_2\) if they are
equal up to deletion of \(\tau\) events. This means that an infinite silent trace is
similar to any trace, consistent with termination insensitivity.

A machine step function
\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
is labeled by an ordered list of security-relevant operations and an event.
We lift it into the security semantics by adding a security context to the
state, giving us a combined step function
\(\mach,\context \stepstounder{\overline{\psi},\obs} \mach',\context' \in
\MACHS \times \CONTEXTS \rightarrow \MACHS \times \CONTEXT \times \mathit{list} ~ \Psi \times \OBSS\).
We will define \(\stepstounder{}\) informally in the next section, and formally
in \cref{sec:props}.

@@

This is a strong threat model, but it does omit some important aspects of
security enforcement: In particular, hardware and timing attacks are out of scope,
and our properties are termination insensitive as a result of how failstops are
modelled (see~\ref{??}).

%\section{Calls, Returns, and Private Allocations}
\section{Properties by Example}
\label{sec:example}

@@

{\tt main} takes a secret as its argument and initializes a variable to contain
potentially sensitive data. We assume that the only observable events are stores
to a special address, called {\tt out}.
{\tt main} calls another function {\tt f},
and afterward may decide to output its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not---instead it will output the results of {\tt f}.

@@

The security semantics for this program is based
on the security-relevant events noted in the right columns of Figures~\ref{fig:main}
and~\ref{fig:f}, namely execution of instructions that allocate or deallocate space,
make a call, or make a return.
Figure \ref{fig:exec1} shows how the corresponding security context
develops over the first few steps of the program. We assume execution begins
at the start of {\tt main} (\(\PCname = 0\)) and that the stack pointer is initially at
address 1000.
\(\xrightarrow{\overline{\psi}}\) transitions are numbered and
written \(\downarrow \overline{\psi}\) in this diagram; events are omitted.

@@

While it would be simple to define integrity as ``all sealed elements retain their
values after the call,'' this would be stricter than necessary. Suppose that
a callee overwrites some data of its caller, but the caller never accesses that data.
By sheer luck, the callee has managed not to impact the observable execution beyond that one write.
Luck, or enforcement: Lazy Tagging and Clearing, from Roessler and Dehon, works by
failstopping any time a caller accesses callee-tainted data, so that out-of-frame
writes should never impact execution.

@@

In place of a predicate we define \(\conf\) as a binary relation on states,
which holds on eventual return states \(\mach'\) and \(\nach'\)
iff all relevant elements that changed between \(\mach\) and \(\mach'\) or \(\nach\) and \(\nach'\)
are the same in \(\mach\) and \(\nach\). We then define
return-time confidentiality as holding if \(\conf\) holds on the matching return states
from \(\mach\) and \(\nach\), respectively.

@@

\section{Machine and Security Semantics, Formally}
\label{sec:props}

@@

Recall from \cref{sec:ideas} that a machine is a map from state elements
to values, with a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
The step function is total---while some enforcement mechanisms may cause a machine
to failstop, we model this as the machine silently diverging by remaining in the
same state.

@@

\subsection{Provenance, Capabilities, and Protecting Objects}

@@

%\paragraph*{Generators}
%
% Generation by
%execution receives as an input a partially instantiated machine state
%and attempts to generate an instruction (or a sequence of instructions
%such as a blessed sequence) that makes sense locally (e.g., jumps go
%to a potentially valid code location and loads read from a
%potentially valid stack location). Then we step the machine and repeat
%the process until we generate or execute some target number of
%instructions, or reach a point where the machine cannot step
%any more.
%
%We extend this technique to keep track of the control flow behavior of
%the program being generated: each time a call or return sequence is
%generated, we ensure that the appropriate policy tags and code
%annotations are set for the entry or return points.% At the same time,
%%we allow the generation to sometimes relax those constraints,
%%introducing potentially ill-formed flows: this causes our programs
%%to failstop when executed in conjunction with the policy monitor,
%%but also allows for revealing errors in our setup.\apt{??}
%
We need to extend typical testing schemes further to handle the nested
nature of confidentiality: rather than just generating two
initial machines that are variants of one another and letting them
execute to test for noninterference, we generate a new variant
{\em every time a call is made} and check confidentiality for the
subtrace produced from that variant state until its corresponding
return. As a result, a ``single'' confidentiality test compactly
checks multiple nested calls.

@@

We write that execution from a state produces an observation trace
\(\mach,\context \hookrightarrow \obsT\) as follows, coinductively:

\judgmenttwo{\((\mach,\context) \stepstounder{} (\mach',\context',\obs)\)}
            {\((\mach',\context') \hookrightarrow \obsT\)}
            {\((\mach,\context) \hookrightarrow \notfinished{\obs}{\obsT}\)}

We define another relation that takes a trace until we have returned from the
active principal.
We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
\(d\) is the depth of the current call.

\judgment{\(|\sigma| < d\)}
         {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \tau\)}

\judgmenttwobrlong{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
                  {\(|\sigma| \geq d\)}
                  {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
                  {\(d \downarrow (\mach,(V,\sigma)) \hookrightarrow \notfinished{\obs}{\obsT}\)}

\paragraph*{Observational Similarity}

We say that two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
is the same. That is, we compare up to deletion of \(\tau\) events.

\definition An element set \(\components\) in state \((\mach,\context)\) contains
irrelevant values, written \((\mach,\context) \parallel \components\), if for all
\(\nach\) such that \(\mach \approx_{\components} \nach\), if
\((\mach,\context) \hookrightarrow \obsT\) and
\((\nach,\context) \hookrightarrow \obsT'\), then
\(\obsT \eqsim \obsT'\).

\judgmenttwo[Returned]
            {\(|\sigma| < d\)}
            {\(P ~ (\mach,(V,\sigma))\)}
            {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}

\judgmenttwobrlong[Returned]
            {\(|\sigma_1| < d\)}
            {\(|\sigma_2| < d\)}
            {\((\mach_1,(V_1,\sigma_1)) ~ R ~ (\mach_2,(V_2,\sigma_2))\)}
            {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}

\definition We define a correct return from \(\mach\) as:
%
\[\ret(\mach) = \{(\mach',\context') \mid \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}\]

\definition
A system enjoys  well-bracketed control flow if, whenever
\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) where
\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\),
\(|\sigma| \uparrow \ret(\mach)\) holds on \((\mach',(V,\sigma))\).

\paragraph*{Integrity}

\definition Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
such that \(\mach[\component] \not = \mach'[\component]\).

\definition We define \(\intProp(\mach,V)\), to hold on a state \((\mach',\context')\) when
\(\{\component \mid V ~ \component = \sealed\} \cap \Delta(\mach,\mach')\)
is irrelevant in \((\mach', \context')\).

\definition A system enjoys {\it stack integrity} if, whenever
\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) with
\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \bar{\psi}\),
we have:
\[(|\sigma'| \uparrow \intProp(\mach',V)) ~ (\mach',(V,\sigma))\]

\definition Let \((\mach,\context)\) be a state with \(\context = (V,\sigma)\), and
\(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\).
Then \((\mach,\context)\) enjoys {\it internal confidentiality} with respect to \(\components\)
if, for any \(\nach\) that is a \(\components\)-variant of \(\mach\), we can take
\(|\sigma| \uparrow (\mach,\context) \hookrightarrow \obsT\) and
\(|\sigma| \uparrow (\nach,\context) \hookrightarrow \obsT'\) and have that
\(\obsT \simeq \obsT'\).

\definition The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
This treats
\(\mach\) and \(\nach\) as the initial states and
\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
set of elements that were corrupted between them.

\definition We define the relation \(\conf(\mach,\nach)\) to hold on a pair of states
\((\mach',\context)\) and \((\nach',\context')\) when
\(\bar{\Diamond}(\mach,\nach,\mach',\nach')\) is irrelevant in \((\mach',\context)\).

\definition A state \((\mach,(V,\sigma))\) enjoys {\it return-time confidentiality}
if, for \(\components = \{\component \mid V ~ \component \in \{\sealed, \unsealed\}\}\)
and any \(\nach\) that is a \(\components\)-variant of \(\mach\), we have:
\[(\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \conf(\mach,\nach) ~ (\nach,(V,\sigma))\]

\definition A system enjoys {\it caller confidentiality} if, whenever
\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',\context',\obs)\)
with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
\((\mach',\context')\) enjoys both internal and return-time confidentiality.

\definition We define \(\cconf(\mach,V)\), to hold on a state \((\mach',\context')\) when
\(\{\component \mid V ~ \component = \unsealed\} \cap \Delta(\mach,\mach')\)
is irrelevant in \((\mach', \context')\).

\definition A system enjoys {\it callee confidentiality} if, whenever
\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\)
with \(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
% or \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
we have:
\[(|\sigma| \uparrow \cconf(\mach',V)) ~ (\mach',\context')\]

>>>>

Stack safety is challenging to specify in full generality, because the stack is fundamentally
an architecture-specific construct. (Contrast Acevedo de Amorim et al.'s \cite{} work on heap safety,
which is broadly consistent across architectures and language semantics, and therefore amenable
to a very general approach.)
The desired behavior of the stack varies based on ISA, calling convention,
and the features of whichever programming language originated the code.
For instance, the current state-of-the-art stack-safety specification from Georges et al.
\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
a callee should always return to its caller. But, in the presence of tail-call elimination, a
return from a callee to a distant ancestor is perfectly reasonable. Any addition of new features
must be accompanied by tweaks to the model.

@@

Security properties tend to be conservative, identifying violations that would be
harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
and of compatibility with existing codebases. In a crucial motivating example for this work,
Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
A set of software-defined rules, termed a micro-policy, determines the tags for operation
results based on the tags of their inputs---or, if the input tags are in a dangerous
combination, it may halt the machine (a ``failstop.'')

Roessler and DeHon find that the intuitive way to enforce stack safety in this setting
is inefficient when allocating and clearing stack frames, because each word must be initialized
with a tag corresponding to the (depth of the) current function activation. They propose
{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
when a function writes to memory. This means that such writes cannot be protected. Instead
the micro-policy enforces that memory be read by the same fuction that wrote it. This
lazy enforcement is incompatible with existing models of stack safety,
because an attacker can modify another function's data. But the owner of
the data will never see the changes, so it is secure in the sense that the attacker is
unsuccessful at changing program behavior (except, perhaps, by causing a fault.) We aim to elaborate
that sense, and to test whether the micro-policy fulfills it.

This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
machine states that make claims about the future observable behavior of execution from those states.
That is not to say that our properties are specialized for lazy micro-policies. It also supports
normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, our
observational properties are even weaker than necessary to support Roessler and Dehon's micro-policy.

@@

The paper is structured as follows. In \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. Then in \cref{sec:machine} we will introduce the machine model that we test, and give
its {\em security semantics}, described in brief below. We formalize the definitions of
stack safety that derive from the security semantics in \cref{sec:props}.
\Cref{sec:testing} describes the testing framework, and
\cref{sec:relwork,sec:future} discuss related and future work.
The remainder of this section introduces ideas that will be important throughout.

\paragraph*{Security Semantics and Property Structure}

@@

The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. Those that we explore in detail in \cref{sec:example}
are shown in white: calls, returns, and the allocation and deallocation of local memory.
In \cref{sec:machine}), we cover additional operations add parameters to the examples', to support
various forms of memory sharing and tail-call elimination. These are shown in light gray.
We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
based on pointer provenance in \cref{app:ptr}, represented by the dark grey
operations, but we do not test this.

@@

\begin{table}
  \begin{tabular}{| l | l |}
    \hline
    \(\psi \in \Psi\) & Parameters \\
    \hline
    \example \(\mathbf{call}\) & Target address, argument registers \\
    \testing & Stack argument offset \& size \\
    \example \(\mathbf{return}\) & \\
    \example \(\mathbf{alloc}\) & Offset \& size \\
    \testing & Public flag \\
    \example \(\mathbf{dealloc}\) & Offset \& size \\
    \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
    \hline
    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
    \hline
    \theory \(\mathbf{promote}\) & Register, offset \& size \\
    \theory \(\mathbf{propagate}\) & Source register/address \\
    \theory & Destination register/address \\
    \theory \(\mathbf{clear}\) & Target register/address \\
    \hline
  \end{tabular}
  \caption{Security-relevant Operations and their Parameters}
  \label{tab:psi}
\end{table}

@@

The security principals in our account of stack safety are function
activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
it may be private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)).

\paragraph*{The Properties and their Hierarchy}

We have five different properties to introduce:

\begin{itemize}
\item Well=bracketed Control Flow (\(WBCF\))
\item Caller Integrity (\(CLRI\))
\item Caller Confidentiality (\(CLRC\))
\item Callee Confidentiality (\(CLEC\))
\item Callee Integrity (\(CLEI\))
\end{itemize}

The properties form a rough hierarchy; although they are technically separate and need
not coexist in any particular combination, some are likely to take priority in terms of their
importance to security. Almost all stack-safety enforcement mechanisms will try to enforce \(WBCF\),
as control-flow attacks such as ROP are a major vulnerability that can be targeted via the stack.
Many also cover some degree of (caller) integrity. [TODO: more hierarchy].

\paragraph*{Events}

@@

@@

This is a strong threat model, but it does omit some important aspects of
security enforcement: In particular, hardware and timing attacks are out of scope.

\section{Properties by Example}
\label{sec:example}

@@

{\tt main} takes a secret as its argument and initializes a variable to contain
potentially sensitive data. We define an observable event to be a store
to a special address, called {\tt out}.
{\tt main} calls another function {\tt f},
and afterward may decide to output its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not---instead it will output the results of {\tt f}.

@@

The security semantics for this program is based
on the security-relevant events noted in the right columns of Figures~\ref{fig:main}
and~\ref{fig:f}, namely execution of instructions that allocate or deallocate space,
make a call, or make a return.
Figure \ref{fig:exec1} shows how the corresponding security context
develops over the first few steps of the program, beginning with
the start of {\tt main} (\(\PCname = 0\)) and with the stack pointer at address 1000.
State transitions are numbered and labeled with a list of security operations, written
\(\downarrow \overline{\psi}\) between steps.

@@

While it would be simple to define integrity as ``all sealed elements retain their
values after the call,'' this would be stricter than necessary. Suppose that
a callee overwrites some data of its caller, but the caller never accesses that data
(or only does so after re-initializing it.) This would be harmless, with the callee
essentially using the caller's memory as scratch space, but the caller never seeing any change.

@@

In place of a predicate we define \(\confProp\) as a binary relation on states,
which holds on eventual return states \(\mach'\) and \(\nach'\)
iff all relevant elements that changed between \(\mach\) and \(\mach'\) or \(\nach\) and \(\nach'\)
are the same in \(\mach\) and \(\nach\). We then define
return-time confidentiality as holding if \(\confProp\) holds on the matching return states
from \(\mach\) and \(\nach\), respectively.

@@

\section{Machine and Security Semantics, Formally}
\label{sec:machine}

@@

A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
\emph{values},  with a step function \(\mach \xrightarrow{\bar{\psi},\obs} \mach'\).
Each value \(v\) contains a \emph{payload} word, written \(|v|\).
The details of value structure depend on the specific machine being modelled;
intuitively, the payload represents the part of the value that is relevant to
the behavior of the basic machine, while the rest of the value may contain
information relevant to a hardware enforcement mechanism (such as a tag).
We write \(\mach[\component]\) to denote the value of \(\mach\) at
\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).

Security classes are ranged over by \(C\).
For some security class \(C\), we write \(C(V)\)
to denote the set of elements \(\component\) such that \(V ~ \component = C\).
The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.

@@

\section{Formal properties}
\label{sec:props}

Now we are ready to introduce the formal properties that define correctly enforced
security semantics. First, we must give a few prelimaries: the generation and comparison
of traces, variant states, and ``on-return'' assertions.

We write the trace of execution from a state up to the first
state below depth \(d\) via the \(\hookrightarrow\) operator, defined coinductively.
When \(d = 0\), the trace will always be infinite. In this case we
omit \(d\) and just write \(\hookrightarrow s\).

\judgment{\(|\sigma| < d\)}
         {\(d \hookrightarrow (\mach,(V,\sigma)) = \tau\)}

\judgmentthree{\((\mach,(V,\sigma)) \stepstounder{} (\mach',\context',\obs)\)}
              {\(|\sigma| \geq d\)}
              {\(d \hookrightarrow (\mach',\context') = \obsT\)}
              {\(d \hookrightarrow (\mach,(V,\sigma)) = \notfinished{\obs}{\obsT}\)}

Two event traces $\obsT_1$ and $\obsT_2$ are {\em similar},
written \(\obsT_1 \eqsim \obsT_2\), if the sequence of non-silent events
is the same. That is, we compare up to deletion of \(\tau\) events.

\definition An element set \(\components\) in state \((\mach,\context)\) contains
irrelevant values, written \((\mach,\context) \parallel \components\), if for all
\(\nach\) such that \(\mach \approx_{\components} \nach\),
\(\hookrightarrow (\mach,\context)  \eqsim \hookrightarrow (\nach,\context)\).

\judgmenttwo[Returned]
            {\(|\sigma| < d\)}
            {\(P ~ \mach\)}
            {\((d \uparrow P) ~ (\mach, (V,\sigma))\)}

\judgmentthree[Returned]
              {\(|\sigma_1| < d\)}
              {\(|\sigma_2| < d\)}
              {\(\mach_1 ~ R ~ \mach_2\)}
              {\((\mach_1,(V_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(V_2,\sigma_2))\)}

[TODO: back-propagate abbreviations used here]

Finally, with the preliminaries out of the way, we can define our core properties. We write
``when \(s\) is called'' as a shorthand for quantifying over all \(s\) reachable in any execution
such that \(s \stepstounder{\overline{\psi}} s'\) and
\(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\).

Let \(\Delta(\mach,\mach')\) be the set of elements \(\component\)
such that \(\mach[\component] \not = \mach'[\component]\).
The {\em corrupted set} \(\bar{\Diamond}(\mach,\mach',\nach,\nach')\)
is the set \((\Delta(\mach,\mach') \cup \Delta(\nach,\nach')) \cap \Delta(\mach',\nach')\).
This treats \(\mach\) and \(\nach\) as the initial states and
\(\mach'\) and \(\nach'\) as their corresponding final states, and is the
set of elements that were corrupted between them.

\begin{table*}
  \begin{align*}
    \mathit{WBCF} \triangleq & |\sigma| \uparrow \ret
    && \textnormal{ where } \ret ~ \mach' \textnormal{ iff }
    \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]+4\}
    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called} \\
    %
    \mathit{ClrI} \triangleq & (|\sigma| \uparrow \intProp) ~ (\mach,(V,\sigma))
    && \textnormal{ where } \intProp ~ \mach' \textnormal{ iff }
    \mach' \parallel \sealed(V) \cap \Delta(\mach,\mach')
    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is called} \\
    %
    \mathit{ClrC} \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\sealed(V)} \nach, &&
    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
    & |\sigma| \hookrightarrow (\mach,(V,\sigma)) \simeq |\sigma| \hookrightarrow (\nach,(V,\sigma)) \\
    & \textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))
    && \textnormal{ where } (\mach' ~ \confProp ~ \nach') \textnormal{ iff }
    \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach') \\
    %
    \mathit{CleC} \triangleq & (|\sigma| \uparrow \cconfProp) ~ (\mach,(V,\sigma))
    && \textnormal{ where } \cconfProp ~ \mach' \textnormal{ iff }
    \mach' \parallel \unsealed(V) \cap \Delta(\mach,\mach')
    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
    %
    \mathit{CleI} \triangleq & \forall \nach \textnormal{ s.t. } \mach \approx_{\unsealed(V)} \nach, &&
    && \textnormal{ when } (\mach,(V,\sigma)) \textnormal{ is (tail)called} \\
    & |\sigma| \hookrightarrow (\mach,(V,\sigma)) \simeq |\sigma| \hookrightarrow (\nach,(V,\sigma)) \\
    & \textnormal{ and } (\mach,(V,\sigma)) ~ (|\sigma| \Uparrow \confProp) ~ (\nach,(V,\sigma))
    && \textnormal{ where } (\mach' ~ \cintProp ~ \nach') \textnormal{ iff }
    \mach' \parallel \bar{\Diamond}(\mach,\nach,\mach',\nach') \\
  \end{align*}
  \caption{Properties}
  \label{tbl:props}
\end{table*}

@@

%\paragraph*{Generators}
%
% Generation by
%execution receives as an input a partially instantiated machine state
%and attempts to generate an instruction (or a sequence of instructions
%such as a blessed sequence) that makes sense locally (e.g., jumps go
%to a potentially valid code location and loads read from a
%potentially valid stack location). Then we step the machine and repeat
%the process until we generate or execute some target number of
%instructions, or reach a point where the machine cannot step
%any more.
%
%We extend this technique to keep track of the control flow behavior of
%the program being generated: each time a call or return sequence is
%generated, we ensure that the appropriate policy tags and code
%annotations are set for the entry or return points.% At the same time,
%%we allow the generation to sometimes relax constraints,
%%introducing potentially ill-formed flows: this causes our programs
%%to failstop when executed in conjunction with the policy monitor,
%%but also allows for revealing errors in our setup.\apt{??}
%
We need to extend typical testing schemes further to handle the nested
nature of confidentiality: rather than just generating two
initial machines that are variants of one another and letting them
execute to test for noninterference, we generate a new variant
{\em every time a call is made} and check confidentiality for the
subtrace produced from that variant state until its corresponding
return. As a result, a ``single'' confidentiality test compactly
checks multiple nested calls.

@@

\subsection{Provenance, Capabilities, and Protecting Objects}
\label{app:ptr}
